On the dynamic assessment of the Local-Maximum Entropy Material Point Method through an Explicit Predictor-Corrector Scheme

### 2020-09-11 LME-MPM.tex
Nueva cita

The aforementioned techniques are devoted to mitigate the
``grid crossing'' error. Nevertheless, in the presence of shock waves, spurious
numerical noises \textcolor{blue}{might} appear despite of the employment of these techniques~\cite{Tran2019e}. These numerical inaccuracies, also known
as wiggles, arise due to inaccuracies in the time discretisation technique.
A simple approach to face those spurious noises lies on the addition of nonphysical damping sources to the equilibrium equations. This
approach has been widely employed in this and many other numerical
techniques. To avoid introducing this nonphysical sources, many
researchers have proposed alternative time integration
schemes which reduce the presence of high frequency noises by
filtering them or increasing somehow the accuracy. One of the first attempts was the proposal of an implicit time integration scheme by Guilkey \& Weiss \cite{Guilkey_2003}. More recently Wang {\it et al.}
\cite{Wang_2016} mitigated these spurious noises by adding a non-viscous
damping to the linear momentum balance equation, and later Charlton
{\it et al.} \cite{Charlton_2017} extended this scheme to the
\acrshort{gimp} approach introducing the \acrfull{igimp}
method. However, the local damping introduced by 
\cite{Wang_2016} can totally over-damp the solution in time-dependent
simulations such as in consolidation process. In a recent publication,
Kularathna \& Soga \cite{Soga_2017} studied an implicit treatment of the pressure in MPM algorithm to simulate material incompressibility
avoiding artificial pressure oscillations by applying
Chorin\textquotesingle s
projection. Within the explicit time integration schemes, Lu {\it et al.}\cite{LU_2018} introduced the time-discontinuous Galerkin method to control the spurious noises
propagation, and later Tran \& Solowski~\cite{Tran2019e}
proposed a generalised-$\alpha$ scheme for \acrshort{mpm} with
promising results but at the expense of increasing the computational
effort. In this paper a less time consuming and high efficient explicit predictor-corrector integration method is
proposed. It consists of an accommodation of the traditional \acrfull{npc} scheme, widely employed in Finite Element methods. \textcolor{blue}{This method
has already been chosen in ~\cite{Navas2018a} among other suitable alternatives} as those proposed
by Wilson {\it et al.} (1972)~\cite{Wilson1972} or Chung \& Hulbert
(1993)~\cite{Geranlized_alpha_1993} because its simplicity and good
performance dealing with solid dynamic problems under a meshfree
framework.  
\textcolor{blue}{Other well known source of energy dissipation and numerical noise is the presence of a non-trivial null space of the linear operator that maps particles values onto nodal values. It was earlier identified by Brackbill~\cite{BRACKBILL1988469} as ringing instability in the \acrfull{pic} method. A recent development in the field of particle methods is a null-space filter engineered by Gritton \&
Berzins~\cite{Gritton2017} which overcome this limitation but still introduce unwelcome damping.  Later, Hammerquist \& Nairn~\cite{HAMMERQUIST2017724} introduced the XPIC as a parametric extension of the aforementioned research with excellent results.}

@@

\textcolor{blue}{The following test is intended to validate the ability of the discussed time integration schemes in order to reproduce the wave propagation in 2D. The impact of an elastic square plate against a vertical wall is assessed. The square plate dimensions are 50 x 50 m and their elastic parameters are:
\begin{itemize} 
\item  Initial density : $20\ kg/m^3$
\item  Poisson ratio : $0.3$
\item  Elastic modulus : $0.1\ MPa$
\end{itemize}
}
\textcolor{blue}{The background set of nodes is composed by a cartesian grid with an uniform spatial discretisation of 1 m. Four material points per cell are considered. The Courant number is fixed to 0.1 during the whole simulation and $\text{LME}_{3.0}$ shape functions are employed. A horizontal velocity of the square plate of $\text{V}_0 = 1.0 m/s$ until the shock against the obstacle, which has been modelled as a set of nodes with prescribed velocity equal to zero. Stress waves are produced by the aforementioned shock what causes the square to bounce back. Both \acrshort{npc} and \acrshort{fe} schemes are compared. Spatial distribution of the velocity field inside the square plate obtained with both schemes at different times are given in figure ~\ref{fig:Magnitude_velocity_impact_square}}. 
%%%%%%%%%%%%%%%%
\begin{figure}
%\sidecaption
  \centering
  \resizebox{\hsize}{!}{
\includegraphics[scale=1]{Figures/Velocity_FE_vs_NPC_square.png} 
  }
  \caption{Spatial distribution of the magnitude of the velocity field in an elastic square plate impacted on the right edge. The top row depicts the \acrshort{npc} results meanwhile the bottom row represents the \acrshort{fe} ones at different instants.}
  \label{fig:Magnitude_velocity_impact_square}
\end{figure}
%%%%%%%%%%%%%%%% 
\textcolor{blue}{At early times, \acrshort{npc} and \acrshort{fe} behvae in a similar manner. However, at later times, the \acrshort{fe} scheme introduces significant numerical noise compared with the \acrshort{npc} one. At the end of the simulation, the filtering process when \acrshort{npc} is used leads to a negligible loss of mechanical energy (about 0.1\%).}

>>>>

The aforementioned techniques are devoted to mitigate the
``grid crossing'' error. Nevertheless, in the presence of shock waves, spurious
numerical noises \textcolor{blue}{might} appear despite of the employment of these techniques~\cite{Tran2019e}. These numerical inaccuracies, also known
as wiggles, arise due to inaccuracies in the time discretisation technique.
A simple approach to face those spurious noises lies on the addition of nonphysical damping sources to the equilibrium equations. This
approach has been widely employed in this and many other numerical
techniques. To avoid introducing this nonphysical sources, many
researchers have proposed alternative time integration
schemes which reduce the presence of high frequency noises by
filtering them or increasing somehow the accuracy. One of the first attempts was the proposal of an implicit time integration scheme by Guilkey \& Weiss \cite{Guilkey_2003}. More recently Wang {\it et al.}
\cite{Wang_2016} mitigated these spurious noises by adding a non-viscous
damping to the linear momentum balance equation, and later Charlton
{\it et al.} \cite{Charlton_2017} extended this scheme to the
\acrshort{gimp} approach introducing the \acrfull{igimp}
method. However, the local damping introduced by 
\cite{Wang_2016} can totally over-damp the solution in time-dependent
simulations such as in consolidation process. In a recent publication,
Kularathna \& Soga \cite{Soga_2017} studied an implicit treatment of the pressure in MPM algorithm to simulate material incompressibility
avoiding artificial pressure oscillations by applying
Chorin\textquotesingle s
projection. Within the explicit time integration schemes, Lu {\it et al.}\cite{LU_2018} introduced the time-discontinuous Galerkin method to control the spurious noises
propagation, and later Tran \& Solowski~\cite{Tran2019e}
proposed a generalised-$\alpha$ scheme for \acrshort{mpm} with
promising results but at the expense of increasing the computational
effort. In this paper a less time consuming and high efficient explicit predictor-corrector integration method is
proposed. It consists of an accommodation of the traditional \acrfull{npc} scheme, widely employed in Finite Element methods. \textcolor{blue}{This method
has already been chosen in ~\cite{Navas2018a} among other suitable alternatives} as those proposed
by Wilson {\it et al.} (1972)~\cite{Wilson1972} or Chung \& Hulbert
(1993)~\cite{Geranlized_alpha_1993} because its simplicity and good
performance dealing with solid dynamic problems under a meshfree
framework.  
\textcolor{blue}{Other well known source of energy dissipation and numerical noise is the presence of a non-trivial null space of the linear operator that maps particles values onto nodal values. It was earlier identified by Brackbill (1984)~\cite{BRACKBILL1988469} as ringing instability in the \acrfull{pic} method. A recent development in the field of particle methods is a null-space filter engineered by Gritton \&
Berzins (2017)~\cite{Gritton2017} which overcome this limitation but still introduce unwelcome damping.  Later, Hammerquist \& Nairn (2017)~\cite{HAMMERQUIST2017724} introduced the XPIC as a parametric extension of the aforementioned research with excellent results.}

@@

\textcolor{blue}{This benchmark has been conveniently adapted from the one proposed by  Hammerquist \& Nairn (2017)~\cite{HAMMERQUIST2017724}, it is intended to validate the ability of the discussed time integration schemes in order to reproduce the wave propagation in 2D. The impact of an elastic square plate against a vertical wall is assessed. The square plate dimensions are 50 x 50 m and their elastic parameters are:
\begin{itemize} 
\item  Initial density : $20\ kg/m^3$
\item  Poisson ratio : $0.3$
\item  Elastic modulus : $0.1\ MPa$
\end{itemize}
}
\textcolor{blue}{The background set of nodes is composed by a cartesian grid with an uniform spatial discretisation of 1 m. Four material points per cell are considered. The Courant number is fixed to 0.1 during the whole simulation and $\text{LME}_{3.0}$ shape functions are employed. A horizontal velocity of the square plate of $\text{V}_0 = 1.0 m/s$ until the shock against the obstacle, which has been modelled as a set of nodes with prescribed velocity equal to zero. Stress waves are produced by the aforementioned shock what causes the square to bounce back. Both \acrshort{npc} and \acrshort{fe} schemes are compared. Spatial distribution of the velocity field inside the square plate obtained with both schemes at different times are given in figure ~\ref{fig:Magnitude_velocity_impact_square}}. 
%%%%%%%%%%%%%%%%
\begin{figure}
%\sidecaption
  \centering
  \resizebox{\hsize}{!}{
\includegraphics[scale=1]{Figures/Velocity_FE_vs_NPC_square.png} 
  }
  \caption{Spatial distribution of the magnitude of the velocity field in an elastic square plate impacted on the right edge. The top row depicts the \acrshort{npc} results meanwhile the bottom row represents the \acrshort{fe} ones at different instants.}
  \label{fig:Magnitude_velocity_impact_square}
\end{figure}
%%%%%%%%%%%%%%%% 
\textcolor{blue}{At early times, \acrshort{npc} and \acrshort{fe} behvae in a similar manner. However, at later times, the \acrshort{fe} scheme introduces significant numerical noise compared with the \acrshort{npc} one. At the end of the simulation, the filtering process when \acrshort{npc} is used leads to a negligible loss of mechanical energy (about 0.1\%).}
